|

### 2020-04-19 paper.Rmd
Australian spelling.

For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed. The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$- dimensional data matrix $\mathbf{X}$. We denote by $\mathbf{A}$, a $p \times d$-dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$. The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

@@

The orthogonal distance of the data points from the current projection plane is calculated as, 
\begin{equation}
\tilde{v}_i^2 = ||\mathbf{x}_i'||^2,
\end{equation}
\noindent where 
\begin{equation}
\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2
\end{equation}
\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$. \textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centered at the origin. Generally we want to centre the plane at the data mean, and this is achieved by centering the data distribution at the origin by subtracting its mean. Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} The quantity $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane. This can be generalized for $d>2$.

The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$. \textcolor{red}{While all projected points may be drawn in the display, point size and/or color can be used to highlight the points inside the slice. This is illustrated in Figure \ref{fig:renderFunction}.} In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$. We could use this single direction to slice systematically from one side of the space to the other, as illustrated in Figure \ref{fig:diagrams}. It is also possible to think of front and back; some points are closer to the viewer, and some are far, and depth cues could be used. \textcolor{red}{However, only $p-d =1$ results in a unique orthogonal direction along which the projection plane can be translated, and in general using distance} from the points to the plane is the simplest approach to generating slices. Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.

>>>>

For the explanation of the slice tour, the actual mechanics of the tour are not important, and only the notion of a projection plane in high dimensions is needed. The data consists of $n$ observations in $p$-dimensions ($n$ points in $\mathbb{R}^p$) and is thus represented as an $n\times p$ dimensional data matrix $\mathbf{X}$. We denote by $\mathbf{A}$, a $p \times d$ dimensional (projection) matrix, the columns of which span a $d$-dimensional subspace of $\mathbb{R}^p$. The $d$-dimensional projection of the data is thus given by $\mathbf{Y} = \mathbf{X}\cdot\mathbf{A}$, producing the $n\times d$ dimensional projected data matrix to be plotted in each frame of the tour display.

@@

The orthogonal distance of the data points from the current projection plane is calculated as, 
\begin{equation}
\tilde{v}_i^2 = ||\mathbf{x}_i'||^2,
\end{equation}
\noindent where 
\begin{equation}
\mathbf{x}_i' = \mathbf{x}_i - (\mathbf{x}_i\cdot \mathbf{a}_1) \mathbf{a}_1 - (\mathbf{x}_i\cdot \mathbf{a}_2) \mathbf{a}_2
\end{equation}
\noindent and $\mathbf{x}_i, i=1,...,n$ is a $p$-dimensional observation in $\mathbf{X}$ and $\mathbf{a}_k, k=1,2 (=d)$ denoting the columns of the projection matrix, $\mathbf{A}=(\mathbf{a}_1, \mathbf{a}_2)$. \textcolor{red}{A priori, the projection matrix $\mathbf{A}$ defines a plane that is centred at the origin. Generally we want to centre the plane at the data mean, and this is achieved by centring the data distribution at the origin by subtracting its mean. Alternatively, we can use a rigid translation to a parallel affine plane that passes through the data mean, as discussed in Section \ref{sec:offcentre}.} The quantity $\mathbf{x}_i'$ can be considered to be a normal from the projection plane to $\mathbf{x}_i$, and then the norm of this vector gives the orthogonal distance between point and plane. This can be generalized for $d>2$.

The distance can then be used to display a slice tour by highlighting \textcolor{red}{the projected} points for which the orthogonal distance is smaller than a selected cutoff value $h$. \textcolor{red}{While all projected points may be drawn in the display, point size and/or colour can be used to highlight the points inside the slice. This is illustrated in Figure \ref{fig:renderFunction}.} In 3D, where for each plane there is only a single orthogonal direction, this defines a flat slice of height $2h$. We could use this single direction to slice systematically from one side of the space to the other, as illustrated in Figure \ref{fig:diagrams}. It is also possible to think of front and back; some points are closer to the viewer, and some are far, and depth cues could be used. \textcolor{red}{However, only $p-d =1$ results in a unique orthogonal direction along which the projection plane can be translated, and in general using distance} from the points to the plane is the simplest approach to generating slices. Using the Euclidean distance results in rotation invariant slicing in the orthogonal space, where a "slice" is spherical in the orthogonal subspace, and has radius $h$.
